\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-05-01 forwards-and-backwards.tex
edits and bib fixes

Specialising the results of 
\citet{KURTZ1999103} to the deterministic setting
provides general conditions under which we
have existence and uniqueness of solutions to~(\ref{eqn:nonlocal_pde}) which
have an $L^2$-density with respect to Lebesgue measure. 
Recall that the Wasserstein metric, defined by 
\[\rho (\nu_1,\nu_2)=\sup\Big\{\Big|\int fd\nu_1-\int fd\nu_2\Big|:\sup_x|f(x
)|\leq 1,|f(x)-f(y)|\leq \|x-y\|\Big\},\]
determines the topology of weak convergence on $\measures$.
We write 
$r(x,\eta) \gamma(x,\eta) \covq(x) = J(x,\eta) J(x,\eta)^T$, and 
$\beta(x,\eta) = r(x,\eta) \gamma(x,\eta) \big(\meanq(x)+2\covq(x)\nabla\log r(x,\eta)\big)$
(quantities that will appear in Proposition~\ref{prop:limiting_construction}).
If $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
that
\begin{equation}
\label{lipschitz for kurtz xiong}
|J (x_1, \nu_1)-J (x_2, \nu_2)|,|\beta (x_1,\nu_1)-\beta 
	(x_2,\nu_2)|, |F(x_1,\nu_1)-F(x_2,\nu_2)|
	%,|b(x_1,\nu_1)-b(x_2,\nu_2)|
	\leq C(\|x_1-x_2\|+\rho (\nu_ 1,\nu_2))
\end{equation}
for some $C>0$,   
the methods of \cite{KURTZ1999103} show that %under the conditions
%of Theorem~\ref{thm:kurtzxiong}, 
if the initial condition $\eta_0$ for our 
population process has
an $L^2$ density, then so does $\eta_t$ for $t>0$. 
Although the necessary estimates (for which we refer to the original paper)
are highly nontrivial, 
the idea of the proof is simple. Take a solution to the
equation and use it to calculate the coefficients $r$, $\gamma$ and $F$ that depend
on local population density. Then $\eta$ solves the 
{\em linear} equation obtained by regarding those values of $r$, $\gamma$ and $F$ as given.
It remains to prove that the solution to the linear equation has a density
which is achieved by obtaining
$L^2$ bounds on its convolution with the heat semigroup at time $\delta$ and letting 
$\delta\to 0$.
We also have the following uniqueness result.
\begin{theorem}[Special case of \cite{KURTZ1999103}, Theorem~3.5]
	\label{thm:kurtzxiong}
Suppose  $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
of~(\ref{lipschitz for kurtz xiong}).
If $\eta_0$ has an $L^2({\mathbb R}^d)$-density, then there 
exists a unique $L^2({\mathbb R}^d)$-valued  
	solution of (\ref{eqn:nonlocal_pde}) in the sense of 
	Definition~\ref{defn:weak_solutions}.
\end{theorem}
\begin{remark}
\cite{KURTZ1999103} considers an infinite system of stochastic differential equations
for the locations and weights of a collection of particles that interact through their 
weighted empirical measure, which is shown to be the unique solution to a stochastic
	PDE. As we shall see through our lookdown representation in 
Section~\ref{sec:lookdown}, the
	solution to our
	deterministic equation can be seen as the empirical measure of a 
	countable number of particles (all with the same weight)
which, in the notation above, evolve according
	to 
	\[X(t)=X(0)+\int_0^t\beta\big(X(s), \eta_s\big)ds
+\int_0^tJ\big(X(s), \eta_s\big)dW(s)\]
	(with an independent Brownian motion $W$ for each particle).
\end{remark}

@@

Now we turn to results about the lookdown process,
first establishing the basic connection between the processes, 
\comment{which processes?}
Proposition~\ref{thm:mmt_application},
and then in the next section, convergence of the lookdown process itself.

\begin{proof}[of Proposition~\ref{thm:mmt_application}]
    This proposition is the content of the Markov Mapping Theorem,
    reproduced from \citet{etheridge/kurtz:2019} as Theorem~\ref{thm:mmt},
    applied to our situation.
    The function $\gamma$ of that theorem is what we have called $\kappa$ above,
    and the kernel $\alpha$ of that theorem
    is the transition function that assigns levels uniformly on $[0, N]$ (in the first case)
    or as a Poisson process with Lesbegue intensity (in the limiting case).
    We need a continuous $\psi^N(\lp)$ such that $|A^N f(\lp)| \le c_f \psi^N(\lp)$
    for all $f$ in the domain of $A^N$ (and similarly a function $\psi$ for $A$).
	\comment{technically need $\psi\geq 1$, so usually just add $1$ to 
	whatever you think $\psi$ should be.} 
    We also need that applying the lookdown generator to a function and averaging over levels
    is equivalent to applying the population process generator to the function
    whose dependence on levels has been averaged out,
    a condition which we precisely state, and verify,
    in Lemmas~\ref{lem:averaged_generators} and~\ref{lem:averaged_generators_limit}
    of the Appendix.

    For finite $N$,
    taking $f(\lp)$ of the form \eqref{eqn:test_functions},
    we can use $\psi^N(\lp) = \langle C (1 + u |F(x, \eta)|), \lp(dx, du) \rangle$
    for an appropriate constant $C$.
    For the scaling limit, recall that the test functions $f$ are of the form
    $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$
    with $g(x, u) = 1$ for $u \ge u_0$,
    and so taking for example the most troublesome term in $Af(\lp)$,
	\comment{should we rephrase to say that the most troublesome term is...
	(because the other one is as above)?}
    we see that since $0 \le f(\lp) / g(x, u) \le 1$ for any $(x, u) \in \lp$,
    \begin{align*} % \label{eqn:part_of_psi}
        \big| f(\lp) \sum_{(x, u) \in \lp} F(x, \eta) u \frac{\partial_u g(x, u)}{g(x, u)} \big|
        & \le
        \| \partial_u g \|_\infty \sum_{(x, u) \in \lp} | F(x, \eta) u \ind_{u \le u_0} | \\
        & \le
        \| \partial_u g \|_\infty e^{u_0} \sum_{(x, u) \in \lp} | F(x, \eta) u e^{-u} | .
    \end{align*}
    The first line would be just what we want,
    except that $\psi(\lp)$ cannot depend on $f$, and hence neither on $u_0$.
    So, the second line provides us with the required bound:
    we absorb $\| \partial_u g \|_\infty e^{u_0}$ into $c_f$
    and take $\psi(\lp) = \langle F(x, \eta) u e^{-u}, \lp(dx, du) \rangle$.
	\comment{again add one - and inside the integral to deal with
	the non-troublesome term}
    % Furthermore, since $F(x, \eta)$ only depends on spatial location, not level,
    % for a given $\lp$ (and hence $\eta = \kappa(\lp)$),
    % $\sup_x |F(x, \eta)|$ is finite.
    % Therefore, we may bound \eqref{eqn:part_of_psi} by
    % $\sup_x |F(x, \eta)| \langle u e^{-u}, \lp \rangle$.
    % To do this we must restrict $\lp$ to lie in the space of measures
    % for which this is finite;
    % however, this is a completely metrizable space
    % \citep[see Condition 2.1 in][]{etheridge/kurtz:2019},
    % and our processes fall in this space (by conditional uniformity).
\end{proof}

@@

Now we turn to the main theorem on convergence of the lookdown process, 
Theorem~\ref{thm:lookdown_convergence}, whose proof
follows a similar pattern to those \comment{`that'?} of convergence for the population processes
in Section~\ref{sec:population_density_proof}.

@@

In addition to tightness of the measure-values processes $\lp^N$,
the bounds used in the proofs below also imply
tightness of the number of lines of descent and the number of births below a fixed level,
and of the motion of individual lines of descent.
In other words, the limiting ``line of descent'' construction
of Section~\ref{sec: individual lines of descent} holds.
	\comment{Is this paragraph really a separate remark that we want to make?}

\begin{proof}[Proof of Theorem~\ref{thm:lookdown_convergence}]
    As in Section~\ref{sec:population_density_proof},
    the theorem will follow from tightness and characterization of the limit.
	\comment{characterisation of limit points (given that we have
	no uniqueness of the limit)?}
    This time, the processes $\lp^N$ take values in $\lpmeasures$,
    the space of locally finite measures on space $\times$ levels.
    (They will in fact be point measures, including the limit,
    but that is a consequence of this theorem.)
    Again, tightness follows from a compact containment condition,
    tightness of one-dimensional distributions,
    and an application of \citet{ethier/kurtz:1986} Theorem~3.9.1.

    Lines of descent can escape to infinite level in finite time, and so
    we endow $\lpmeasures$ with the vague topology ``in the level coordinate'',
    induced by test functions on $\overline{\IR^d} \times [0,\infty)$
    of the form $g(x) h(u)$, where $g \in C_b(\overline{\IR^d})$ is bounded and continuous
    and $h \in C_c([0, \infty))$ is compactly supported.
    This space is completely metrizable,
    and we may choose a countable set of nonnegative, compactly supported $f_k$
    such that a subset $K \subset \lpmeasures$ is relatively compact if and only if
    $\sup_{\lp \in K} \langle f_k, \lp \rangle < \infty$ for each $k$;
    below, Lemma~\ref{lem:compact containment lookdown} proves exactly this, and therefore compact containment.
	\comment{Do we have a reference for this claim about completely
	metrizable and these sets being enough to 
	establish relative compactness? The reader may reasonably
	ask why we had to add a point at infinity to $\IR^d$ and not one
	to $[0,\infty)$. Also, maybe spell out that the support of $f_k$ is compact in
	the product space of $\overline{\IR^d}$ times levels?}

    We require that $\{(F(\lp^N_t))_{t \ge 0}\}_N$
    is tight as a sequence of real-valued c\`adl\`ag processes, for all $F$ in a subset
    of $\bar C(\lpmeasures)$, the bounded, continuous functions on $\lpmeasures$,
    that is dense with respect to uniform convergence on compact subsets.
    Lemma~\ref{lem:test_fn_tightness} shows that $\{\langle f, \lp^N_t \rangle\}_N$ is a tight sequence
    for any $f: \overline{\IR^d} \times [0,\infty) \to \IR$ with compact support
    in the level direction.
    The functions $\lp \mapsto \langle f, \lp \rangle$
    do not form not a dense subset of $\bar C(\lpmeasures)$,
    but they do separate points and vanish nowhere, 
	\comment{I don't understand vanish nowhere - I think it may because
	of the way I am reading the sentence, but it is perfectly possible
	for $\langle f,\xi\rangle$ to be zero.}
	so
    by the Stone-Weierstrass theorem, the algebra they generate is dense
    % in $C_0(\lpmeasures)$, the continuous functions that vanish at infinity
    % (i.e., are less than $\epsilon$ outside of a compact set, for any $\epsilon$).
    % Since $C_0(\lpmeasures)$ is dense
    in $\bar C(\lpmeasures)$ (with respect to uniform convergence on compact subsets).
    So, it suffices to show that tightness for the processes $\langle f, \lp^N_t\rangle$
    extends to finite sums and products of these processes,
    which is shown in Lemma~\ref{lem:product_tightness}.
    The fact that martingale properties are preserved under passage to the limit
    is straightforward, and can be proved in a way analogous to Lemma~\ref{lem:limit_mgale};
    we omit the proof.
    Finally, we must show that the limiting lookdown process $\lp$
    projects to the limiting process $\eta$, i.e., a solution of the martingale
    problem in Theorem~\ref{thm:nonlocal_convergence}.
    Let $N_k \to \infty$ be a sequence along which $\lp^{N_k}$ converges.
    By Theorem~\ref{thm:nonlocal_convergence},
    there is a subsequence $N_{k(j)}$ along which the projected population processes
    $\eta^{N_{k(j)}}$ converge, and the limit solves the martingale problem.
    Thus any limit point of $\lp^N$ projects to a population process $\eta$
    solving the martingale problem of Theorem~\ref{thm:nonlocal_convergence}.
\end{proof}

@@

\begin{lemma}
        \label{lem:h_containment}
    Let $h$ be a positive, continuous, nonincreasing, differentiable function on $[0, \infty)$
    such that
    $\int_0^\infty \int_u^\infty h(v) dv du$,
    $\int_0^\infty u^2 |h'(u)| du$,
    and $\int_0^\infty h(u)^2 du$
    are all finite,
	\comment{Should be a fullstop.}
    Suppose that Assumptions~\ref{def:model_setup} hold,
    and that $\theta/N \to \alpha$ and $\lp_0^N \to \lp_0$ weakly as $N \to \infty$,
    where each $\lp_0^N$ is conditionally uniform given $\eta_0^N$
    in the sense of \eqref{eqn:conditionally_uniform}
    and $\lp_0$ is conditionally Poisson given $\eta_0$
    in the sense of \eqref{eqn:conditionally_poisson}.
    Then for any $T$ there exists a constant $K(T)$ such that for all $M > 0$,
    $$ 
    \limsup_{N \to \infty} \IP\left\{ \sup_{0 \le t \le T} \langle h, \lp^N_t \rangle > M \right\}
    <
    \frac{K(T)}{M} .
    $$
\end{lemma}

We postpone the proof of this Lemma until we have shown how it yields 
\comment{delete `to'} to compact containment.
First, we show that this implies compact containment
\comment{compact containment of what?} for arbitrary
compactly supported functions.

\begin{lemma}
    \label{lem:compact containment hn}
    Suppose $f \in C(\bar \IR^d \times [0, \infty))$ and there is a $u_f$ such that
    $f(x, u) = 0$ for all $x$ and $u \ge u_f$.
	\comment{sounds like $f$ vanishes for all $x$ irrespective of $u$!
	Can we reword? `all $(x,u)$ with $u\geq u_f$' or something?}
    Under the assumptions of Lemma~\ref{lem:h_containment},
    for any $T$ there exists a constant $K(f,T)$ such that for all $M > 0$,
    $$ 
    \limsup_{N \to \infty} \IP\left\{ \sup_{0 \le t \le T} \langle f, \lp^N_t \rangle > M \right\}
    <
    \frac{K(f,T)}{M} .
    $$
\end{lemma}

@@

In other words, the processes $\lp^N$ stay in 
the set
$$
    \left\{
        \lp \in \lpmeasures \;:\;
        \langle f_k, \lp \rangle \le C_k
        \text{ for all } k \ge 1
    \right\} ,
$$
for all $0 \le t \le T$
with uniformly high probability,
a set which is compact if we choose $\{f_k\}$ appropriately.
\comment{Does the reader know how to act `appropriately' in 
making this choice? Maybe refer back to the right bit above where it 
is explained.}
% \comment{
%     Note that we could instead of the vague topology on locally finite measures,
%     take the space of measures with $\langle h, \xi \rangle < \infty$
%     for which $\xi_n \to \xi$ if $\langle f, \xi \rangle < \infty$ for all $f$ for which $f \le c_f h$,
%     which is completely metrizable
%     as in Condition 2.1 of \citet{etheridge/kurtz:2019}.
% }

@@

    Again \comment{Only a very alert reader will remember why we say `again'
    here...} since $\sqrt{x} \le 1 + x$ for $x \ge 0$,
    the Burkholder-Davis-Gundy inequality
    tells us that there is a $C'$ such that
    \begin{align*}
    \IE\left[ \sup_{0 \le t \le T} M^{h}_t \right]
    & \le
        C'\left( 1 + \IE\left[ [M^{h}]_T \right] \right)
    \\ & \le
        C' \left( 1 + \int_0^T \IE\left[
        \langle 2 c_\theta(x, \eta_s^N) \int_u^\infty h(v)^2 dv
        ,
        \lp_s^N(dx, du) \big \rangle
        \right] ds \right) 
    \\ & \le
        C' \left( 1 + 2 C_a \int_0^\infty h(v)^2 dv
        \int_0^T \IE\left[
            \langle 1 , \lp_s^N(dx, du) \big \rangle
        \right] ds \right)  
    \\ & \le
        C_2 e^{C_1 T} ,
    \end{align*}
    for a constant $C_2$ which is finite
    by our assumption that $\int_0^\infty h(v)^2 dv < \infty$.

@@

\begin{lemma}
    \label{lem:test_fn_tightness}
    Let $f$ be a bounded, continuous real-valued function on $\IR^d \times [0, \infty)$
    with uniformly bounded first and second derivatives
    for which there exists a $u_0$ such that if $u > u_0$ then $f(x, u) = 0$.
    Then, the sequence of real-valued processes $(\langle f, \lp_t^N \rangle)_{t \ge 0}$
	are \comment{`is' - the subject of the sentence is the (single) sequence}
	tight in the space of \comment{real-valued? or ${\cal D}_\IR$-valued} 
	c\`adl\`ag processes.
\end{lemma}

>>>>

Specialising the results of 
\citet{kurtz/xiong:1999} to the deterministic setting
provides general conditions under which we
have existence and uniqueness of solutions to~(\ref{eqn:nonlocal_pde}) which
have an $L^2$-density with respect to Lebesgue measure. 
Recall that the Wasserstein metric, defined by 
\[\rho (\nu_1,\nu_2)=\sup\Big\{\Big|\int fd\nu_1-\int fd\nu_2\Big|:\sup_x|f(x
)|\leq 1,|f(x)-f(y)|\leq \|x-y\|\Big\},\]
determines the topology of weak convergence on $\measures$.
We write 
$r(x,\eta) \gamma(x,\eta) \covq(x) = J(x,\eta) J(x,\eta)^T$, and 
$\beta(x,\eta) = r(x,\eta) \gamma(x,\eta) \big(\meanq(x)+2\covq(x)\nabla\log r(x,\eta)\big)$
(quantities that will appear in Proposition~\ref{prop:limiting_construction}).
If $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
that
\begin{equation}
\label{lipschitz for kurtz xiong}
|J (x_1, \nu_1)-J (x_2, \nu_2)|,|\beta (x_1,\nu_1)-\beta 
	(x_2,\nu_2)|, |F(x_1,\nu_1)-F(x_2,\nu_2)|
	%,|b(x_1,\nu_1)-b(x_2,\nu_2)|
	\leq C(\|x_1-x_2\|+\rho (\nu_ 1,\nu_2))
\end{equation}
for some $C>0$,   
the methods of \cite{KURTZ1999103} show that %under the conditions
%of Theorem~\ref{thm:kurtzxiong}, 
if the initial condition $\eta_0$ for our 
population process has
an $L^2$ density, then so does $\eta_t$ for $t>0$. 
Although the necessary estimates (for which we refer to the original paper)
are highly nontrivial, 
the idea of the proof is simple. Take a solution to the
equation and use it to calculate the coefficients $r$, $\gamma$ and $F$ that depend
on local population density. Then $\eta$ solves the 
{\em linear} equation obtained by regarding those values of $r$, $\gamma$ and $F$ as given.
It remains to prove that the solution to the linear equation has a density
which is achieved by obtaining
$L^2$ bounds on its convolution with the heat semigroup at time $\delta$ and letting 
$\delta\to 0$.
We also have the following uniqueness result.
\begin{theorem}[Special case of \cite{KURTZ1999103}, Theorem~3.5]
	\label{thm:kurtzxiong}
Suppose  $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
of~(\ref{lipschitz for kurtz xiong}).
If $\eta_0$ has an $L^2({\mathbb R}^d)$-density, then there 
exists a unique $L^2({\mathbb R}^d)$-valued  
	solution of (\ref{eqn:nonlocal_pde}) in the sense of 
	Definition~\ref{defn:weak_solutions}.
\end{theorem}
\begin{remark}
\cite{KURTZ1999103} considers an infinite system of stochastic differential equations
for the locations and weights of a collection of particles that interact through their 
weighted empirical measure, which is shown to be the unique solution to a stochastic
	PDE. As we shall see through our lookdown representation in 
Section~\ref{sec:lookdown}, the
	solution to our
	deterministic equation can be seen as the empirical measure of a 
	countable number of particles (all with the same weight)
which, in the notation above, evolve according
	to 
	\[X(t)=X(0)+\int_0^t\beta\big(X(s), \eta_s\big)ds
+\int_0^tJ\big(X(s), \eta_s\big)dW(s)\]
	(with an independent Brownian motion $W$ for each particle).
\end{remark}

@@

Now we turn to results about the lookdown process,
first establishing the basic connection between the population process $\eta^N$
and the lookdown process $\lp^N$,
Proposition~\ref{thm:mmt_application},
and then in the next section, convergence of the lookdown process itself.

\begin{proof}[of Proposition~\ref{thm:mmt_application}]
    This proposition is the content of the Markov Mapping Theorem,
    reproduced from \citet{etheridge/kurtz:2019} as Theorem~\ref{thm:mmt},
    applied to our situation.
    The function $\gamma$ of that theorem is what we have called $\kappa$ above,
    and the kernel $\alpha$ of that theorem
    is the transition function that assigns levels uniformly on $[0, N]$ (in the first case)
    or as a Poisson process with Lesbegue intensity (in the limiting case).
    We need a continuous $\psi^N(\lp) \ge 1$ such that $|A^N f(\lp)| \le c_f \psi^N(\lp)$
    for all $f$ in the domain of $A^N$ (and similarly a function $\psi$ for $A$).
    We also need that applying the lookdown generator to a function and averaging over levels
    is equivalent to applying the population process generator to the function
    whose dependence on levels has been averaged out,
    a condition which we precisely state, and verify,
    in Lemmas~\ref{lem:averaged_generators} and~\ref{lem:averaged_generators_limit}
    of the Appendix.

    For finite $N$,
    taking $f(\lp)$ of the form \eqref{eqn:test_functions},
    we can use $\psi^N(\lp) = \langle C (1 + u |F(x, \eta)|), \lp(dx, du) \rangle$
    for an appropriate constant $C$.
    For the scaling limit, recall that the test functions $f$ are of the form
    $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$
    with $g(x, u) = 1$ for $u \ge u_0$,
    and consulting~\eqref{eqn:limiting_lookdown_generator},
    we see that most terms in $Af(\lp)$ can be bounded 
    as above by constant multiples of $\langle 1, \eta \rangle$.
    However, the term involving $F$ is, as usual, more troublesome.
    Since $0 \le f(\lp) / g(x, u) \le 1$ for any $(x, u) \in \lp$,
    \begin{align*} % \label{eqn:part_of_psi}
        \big| f(\lp) \sum_{(x, u) \in \lp} F(x, \eta) u \frac{\partial_u g(x, u)}{g(x, u)} \big|
        & \le
        \| \partial_u g \|_\infty \sum_{(x, u) \in \lp} | F(x, \eta) u \ind_{u \le u_0} | \\
        & \le
        \| \partial_u g \|_\infty e^{u_0} \sum_{(x, u) \in \lp} | F(x, \eta) u e^{-u} | .
    \end{align*}
    The first line would be just what we want,
    except that $\psi(\lp)$ cannot depend on $f$, and hence neither on $u_0$.
    So, the second line provides us with the required bound:
    we absorb $\| \partial_u g \|_\infty e^{u_0}$ into $c_f$
    and take $\psi(\lp) = 1 + \langle 1 + F(x, \eta) u e^{-u}, \lp(dx, du) \rangle$.
    % Furthermore, since $F(x, \eta)$ only depends on spatial location, not level,
    % for a given $\lp$ (and hence $\eta = \kappa(\lp)$),
    % $\sup_x |F(x, \eta)|$ is finite.
    % Therefore, we may bound \eqref{eqn:part_of_psi} by
    % $\sup_x |F(x, \eta)| \langle u e^{-u}, \lp \rangle$.
    % To do this we must restrict $\lp$ to lie in the space of measures
    % for which this is finite;
    % however, this is a completely metrizable space
    % \citep[see Condition 2.1 in][]{etheridge/kurtz:2019},
    % and our processes fall in this space (by conditional uniformity).
\end{proof}

@@

Now we turn to the main theorem on convergence of the lookdown process, 
Theorem~\ref{thm:lookdown_convergence}, whose proof
follows a similar pattern to that of convergence for the population processes
in Section~\ref{sec:population_density_proof}.

@@

In addition to tightness of the measure-values processes $\lp^N$,
the bounds used in the proofs below also imply
tightness of the number of lines of descent and the number of births below a fixed level,
and of the motion of individual lines of descent.
In other words, the limiting ``line of descent'' construction
of Section~\ref{sec: individual lines of descent} holds.
	\comment{Is this paragraph really a separate remark that we want to make?}
    \comment{PETER: maybe? Figuring out how to make it more concrete/rigorous seems daunting, though -
        I'm inclined to remove it.}

\begin{proof}[Proof of Theorem~\ref{thm:lookdown_convergence}]
    As in Section~\ref{sec:population_density_proof},
    the theorem will follow from tightness and characterization of the limit points.
    This time, the processes $\lp^N$ take values in $\lpmeasures$,
    the space of locally finite measures on space $\times$ levels.
    (They will in fact be point measures, including the limit,
    but that is a consequence of this theorem.)
    Again, tightness follows from a compact containment condition,
    tightness of one-dimensional distributions,
    and an application of \citet{ethier/kurtz:1986} Theorem~3.9.1.

    Lines of descent can escape to infinite level in finite time, and so
    we endow $\lpmeasures$ with the vague topology ``in the level coordinate'',
    induced by test functions on $\overline{\IR^d} \times [0,\infty)$
    of the form $g(x) h(u)$, where $g \in C_b(\overline{\IR^d})$ is bounded and continuous
    and $h \in C_c([0, \infty))$ is compactly supported
    (following, e.g., \citet{etheridge/kurtz:2019}, Condition 2.1).
    % PLR: NEW DETAILS
    Several places below we require a dense subset
    of $\bar C(\lpmeasures)$, the bounded, continuous functions on $\lpmeasures$.
    The functions $\lp \mapsto \langle f, \lp \rangle$
    do not form not a dense subset of $\bar C(\lpmeasures)$,
    but they do separate points and vanish nowhere, 
    i.e., for any $\lp_1$ and $\lp_2$ there is an $f$
    with $\langle f, \lp_1\rangle \neq \langle f, \lp_2\rangle$,
    and a $g$ such that $\langle g, \lp_1\rangle \neq 0$.
	Therefore,
    by the Stone-Weierstrass theorem, the algebra they generate is dense
    % in $C_0(\lpmeasures)$, the continuous functions that vanish at infinity
    % (i.e., are less than $\epsilon$ outside of a compact set, for any $\epsilon$).
    % Since $C_0(\lpmeasures)$ is dense
    in $\bar C(\lpmeasures)$ (with respect to uniform convergence on compact subsets).
    % END NEW DETAILS
    Topologized in this way, the space $\lpmeasures$ completely metrizable,
    and we may choose a countable set of nonnegative $f_k$,
    each supported on $\IR^d \times [0,u_k]$ for some $u_k < \infty$,
    such that a subset $K \subset \lpmeasures$ is relatively compact if and only if
    $\sup_{\lp \in K} \langle f_k, \lp \rangle < \infty$ for each $k$.
    (To see this, use Theorem A.2.3 of~\citet{kallenberg1997foundations}
    and the first argument made in the proof of Lemma~\ref{lem:product_tightness}.)
    Below, Lemma~\ref{lem:compact containment lookdown} proves exactly this,
    and therefore compact containment.
    Here we have compactified $\IR^d$ for convenience
    (since it turned out to be straightforward to show that mass does not escape to infinity in space);
    however, we need to use the vague topology ``in the level direction''
    because \emph{levels} may escape to infinity in finite time in the limit.

    For \citet{ethier/kurtz:1986} Theorem~3.9.1
    we require that $\{(F(\lp^N_t))_{t \ge 0}\}_N$
    is tight as a sequence of real-valued c\`adl\`ag processes, for all $F$ in a subset
    of $\bar C(\lpmeasures)$ %, the bounded, continuous functions on $\lpmeasures$,
    that is dense with respect to uniform convergence on compact subsets.
    Lemma~\ref{lem:test_fn_tightness} shows that $\{\langle f, \lp^N_t \rangle\}_N$ is a tight sequence
    for any $f: \overline{\IR^d} \times [0,\infty) \to \IR$ with compact support
    in the level direction.
    % The functions $\lp \mapsto \langle f, \lp \rangle$
    % do not form not a dense subset of $\bar C(\lpmeasures)$,
    % but they do separate points and vanish nowhere, 
    % i.e., for any $\lp_1$ and $\lp_2$ there is an $f$
    % with $\langle f, \lp_1\rangle \neq \langle f, \lp_2\rangle$,
    % and a $g$ such that $\langle g, \lp_1\rangle \neq 0$.
	% Therefore,
    % by the Stone-Weierstrass theorem, the algebra they generate is dense
    % % in $C_0(\lpmeasures)$, the continuous functions that vanish at infinity
    % % (i.e., are less than $\epsilon$ outside of a compact set, for any $\epsilon$).
    % % Since $C_0(\lpmeasures)$ is dense
    % in $\bar C(\lpmeasures)$ (with respect to uniform convergence on compact subsets).  So, 
    Since as above the algebra generated by these are dense in $\bar C(\lpmeasures)$,
    it suffices to show that tightness for the processes $\langle f, \lp^N_t\rangle$
    extends to finite sums and products of these processes,
    which is shown in Lemma~\ref{lem:product_tightness}.
    The fact that martingale properties are preserved under passage to the limit
    is straightforward, and can be proved in a way analogous to Lemma~\ref{lem:limit_mgale};
    we omit the proof.
    Finally, we must show that the limiting lookdown process $\lp$
    projects to the limiting process $\eta$, i.e., a solution of the martingale
    problem in Theorem~\ref{thm:nonlocal_convergence}.
    Let $N_k \to \infty$ be a sequence along which $\lp^{N_k}$ converges.
    By Theorem~\ref{thm:nonlocal_convergence},
    there is a subsequence $N_{k(j)}$ along which the projected population processes
    $\eta^{N_{k(j)}}$ converge, and the limit solves the martingale problem.
    Thus any limit point of $\lp^N$ projects to a population process $\eta$
    solving the martingale problem of Theorem~\ref{thm:nonlocal_convergence}.
\end{proof}

@@

\begin{lemma}
        \label{lem:h_containment}
    Let $h$ be a positive, continuous, nonincreasing, differentiable function on $[0, \infty)$
    such that
    $\int_0^\infty \int_u^\infty h(v) dv du$,
    $\int_0^\infty u^2 |h'(u)| du$,
    and $\int_0^\infty h(u)^2 du$
    are all finite.
    Suppose that Assumptions~\ref{def:model_setup} hold,
    and that $\theta/N \to \alpha$ and $\lp_0^N \to \lp_0$ weakly as $N \to \infty$,
    where each $\lp_0^N$ is conditionally uniform given $\eta_0^N$
    in the sense of \eqref{eqn:conditionally_uniform}
    and $\lp_0$ is conditionally Poisson given $\eta_0$
    in the sense of \eqref{eqn:conditionally_poisson}.
    Then for any $T$ there exists a constant $K(T)$ such that for all $M > 0$,
    $$ 
    \limsup_{N \to \infty} \IP\left\{ \sup_{0 \le t \le T} \langle h, \lp^N_t \rangle > M \right\}
    <
    \frac{K(T)}{M} .
    $$
\end{lemma}

We postpone the proof of this Lemma until we have shown how it yields compact containment.
First, we show that this implies compact containment of the processes $(\langle f, \lp^N_t \rangle)_{0 \le t \le T}$
for arbitrary compactly supported $f$.

\begin{lemma}
    \label{lem:compact containment hn}
    Suppose $f \in C(\bar \IR^d \times [0, \infty))$ and there is a $u_f$ such that
    if $u \ge u_f$ then $\sup_x f(x, u) = 0$.
    Under the assumptions of Lemma~\ref{lem:h_containment},
    for any $T$ there exists a constant $K(f,T)$ such that for all $M > 0$,
    $$ 
    \limsup_{N \to \infty} \IP\left\{ \sup_{0 \le t \le T} \langle f, \lp^N_t \rangle > M \right\}
    <
    \frac{K(f,T)}{M} .
    $$
\end{lemma}

@@

In other words, the processes $\lp^N$ stay in 
the set
$$
    \left\{
        \lp \in \lpmeasures \;:\;
        \langle f_k, \lp \rangle \le C_k
        \text{ for all } k \ge 1
    \right\} ,
$$
for all $0 \le t \le T$
with uniformly high probability,
a set which (as discussed in the proof of Theorem~\ref{thm:lookdown_convergence})
is relatively compact by appropriate choice of $\{f_k\}_{k \ge 1}$.
% \comment{
%     Note that we could instead of the vague topology on locally finite measures,
%     take the space of measures with $\langle h, \xi \rangle < \infty$
%     for which $\xi_n \to \xi$ if $\langle f, \xi \rangle < \infty$ for all $f$ for which $f \le c_f h$,
%     which is completely metrizable
%     as in Condition 2.1 of \citet{etheridge/kurtz:2019}.
% }

@@

    Since $\sqrt{x} \le 1 + x$ for $x \ge 0$,
    the Burkholder-Davis-Gundy inequality
    tells us that there is a $C'$ such that
    \begin{align*}
    \IE\left[ \sup_{0 \le t \le T} M^{h}_t \right]
    & \le
        C'\left( 1 + \IE\left[ [M^{h}]_T \right] \right)
    \\ & \le
        C' \left( 1 + \int_0^T \IE\left[
        \langle 2 c_\theta(x, \eta_s^N) \int_u^\infty h(v)^2 dv
        ,
        \lp_s^N(dx, du) \big \rangle
        \right] ds \right) 
    \\ & \le
        C' \left( 1 + 2 C_a \int_0^\infty h(v)^2 dv
        \int_0^T \IE\left[
            \langle 1 , \lp_s^N(dx, du) \big \rangle
        \right] ds \right)  
    \\ & \le
        C_2 e^{C_1 T} ,
    \end{align*}
    for a constant $C_2$ which is finite
    by our assumption that $\int_0^\infty h(v)^2 dv < \infty$.

@@

\begin{lemma}
    \label{lem:test_fn_tightness}
    Let $f$ be a bounded, continuous real-valued function on $\IR^d \times [0, \infty)$
    with uniformly bounded first and second derivatives
    for which there exists a $u_0$ such that if $u > u_0$ then $f(x, u) = 0$.
    Then, the sequence of real-valued processes $(\langle f, \lp_t^N \rangle)_{t \ge 0}$ for $N \ge 1$
    is tight in ${\cal D}_{[0\infty)}(\IR)$.
\end{lemma}
